---
title: "how to handle high-cardinality predictors for data on museums in the UK"
author: "chad allison"
date: '2023-01-25'
output: github_document
---

### setup

```{r message = F, warning = F}
tictoc::tic()
library(tidyverse) # essential functions
library(tidymodels) # modeling
library(embed) # feature encoding
library(skimr) # data skimming
library(finetune) # model tuning
library(vip) # variable importance
theme_set(theme_minimal())
update_geom_defaults("rect", list(fill = "springgreen4", alpha = 0.75))
```

### data import

```{r}
link = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-11-22/museums.csv"
museums = read_csv(link, col_types = cols())
glimpse(museums)
```

### visualizing `Accrediation` counts (our outcome variable)

```{r}
museums |>
  count(Accreditation) |>
  ggplot(aes(Accreditation, n)) +
  geom_col(aes(fill = Accreditation), alpha = 0.75) +
  geom_text(aes(label = n), vjust = -0.5) +
  scale_fill_manual(values = c("springgreen4", "indianred3")) +
  labs(x = "accreditation status", y = "count",
       title = "accreditation counts; not too much class imbalance") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")
```

### closer look at top subject matters

```{r}
top_subjects = museums |>
  count(Subject_Matter, sort = T) |>
  slice_max(n, n = 6) |>
  pull(Subject_Matter)

museums |>
  filter(Subject_Matter %in% top_subjects) |>
  count(Subject_Matter, Accreditation) |>
  ggplot(aes(Accreditation, n)) +
  geom_col(aes(fill = Accreditation), show.legend = F, alpha = 0.75) +
  scale_fill_manual(values = c("springgreen4", "indianred3")) +
  facet_wrap(vars(Subject_Matter), scales = "free") +
  labs(title = "accreditation counts among top six subjects") +
  theme(plot.title = element_text(hjust = 0.5))
```

### accreditation counts for different museum sizes

```{r}
museums |>
  count(Accreditation, Size) |>
  mutate(Size = factor(Size, levels = c("unknown", "small", "medium", "large", "huge"))) |>
  ggplot(aes(Size, n)) +
  geom_col(aes(fill = Accreditation), position = "dodge", alpha = 0.75) +
  scale_fill_manual(values = c("springgreen4", "indianred3")) +
  annotate("text", x = 4.5, y = 1375, size = 3.5, fontface = "italic",
           label = "note that some of the musuem\nsize variables are imputed") +
  labs(x = "museum size", y = "count", fill = NULL,
       title = "accreditation status by museum size") +
  theme(plot.title = element_text(hjust = 0.5))
```

### governance status and accreditation counts

```{r}
top_gov = museums |>
  count(Governance, sort = T) |>
  slice_max(n, n = 4) |>
  pull(Governance)

museums |>
  filter(Governance %in% top_gov) |>
  count(Governance, Accreditation) |>
  ggplot(aes(Accreditation, n)) +
  geom_col(aes(fill = Accreditation), show.legend = F, alpha = 0.75) +
  scale_fill_manual(values = c("springgreen4", "indianred3")) +
  facet_wrap(vars(Governance), scales = "free_y") +
  labs(x = "accreditation status", y = "count",
       title = "accreditation counts among top governance statuses") +
  theme(plot.title = element_text(hjust = 0.5))
```

### getting data for modeling

```{r}
museums_parsed = museums |>
  select(museum_id, Accreditation, Governance, Size, Subject_Matter,
         Year_opened, Year_closed, Area_Deprivation_index) |>
  mutate(Year_opened = parse_number(Year_opened),
         Closed = ifelse(Year_closed == "9999:9999", "Open", "Closed")) |>
  select(-Year_closed) |>
  na.omit() |>
  mutate_if(is.character, as.factor) |>
  mutate(museum_id = as.character(museum_id))

glimpse(museums_parsed)
```

### data splitting and cross validation

```{r}
set.seed(123)
museum_split = initial_split(museums_parsed, prop = 0.75, strata = Accreditation)
museum_train = training(museum_split)
museum_test = testing(museum_split)

set.seed(234)
museum_folds = vfold_cv(museum_train, strata = Accreditation)
museum_folds
```

### creating recipe while handling high-cardinality

```{r}
museum_rec = recipe(Accreditation ~ ., data = museum_train) |>
  update_role(museum_id, new_role = "id") |>
  step_lencode_glm(Subject_Matter, outcome = vars(Accreditation)) |> # embed
  step_dummy(all_nominal_predictors())

prep(museum_rec) |>
  bake(new_data = NULL) |>
  skim()
```

### looking at the effects of the subject matter encodings

```{r}
prep(museum_rec) |>
  tidy(number = 1) |>
  filter(level == "..new") |>
  mutate(value = round(value, 3))
```

### creating model specification

```{r}
xgb_spec = boost_tree(trees = tune(), min_n = tune(), mtry = tune(), learn_rate = 0.01) |>
  set_engine("xgboost") |>
  set_mode("classification")

xgb_spec
```

### creating workflow

```{r}
xgb_wf = workflow(museum_rec, xgb_spec)
xgb_wf
```

### tuning

```{r}
doParallel::registerDoParallel()
set.seed(345)
xgb_rs = tune_race_anova(xgb_wf,
                resamples = museum_folds,
                grid = 15,
                control = control_race(verbose_elim = T))

xgb_rs
```

### plotting tuning results

```{r}
plot_race(xgb_rs) +
  labs(x = "analysis stage", title = "racing results") +
  theme(plot.title = element_text(hjust = 0.5))
```

### modeling metrics

```{r}
collect_metrics(xgb_rs)
```

### fitting model on all training data, evaluating on testing data

```{r}
xgb_last = xgb_wf |>
  finalize_workflow(select_best(xgb_rs, "accuracy")) |>
  last_fit(museum_split)

xgb_last |>
  collect_metrics()
```

### confusion matrix

```{r}
collect_predictions(xgb_last) |>
  conf_mat(Accreditation, .pred_class) |>
  autoplot(type = "heatmap")
```

### variable importance

```{r}
xgb_last |>
  extract_fit_engine() |>
  vip() +
  annotate("text", x = 5, y = 0.15, size = 3.5, fontface = "italic",
           label = "note that subject matter is important;\nfeature encoding enabled this")
```

### visualizing original data given variable importance

```{r}
museums |>
  mutate(closed = ifelse(Year_closed == "9999:9999", "open", "closed")) |>
  count(Accreditation, closed) |>
  ggplot(aes(closed, n)) +
  geom_col(aes(fill = Accreditation), position = "dodge") +
  geom_text(aes(label = n), vjust = -0.5, position = position_dodge2(width = 0.9), size = 3.5) +
  scale_fill_manual(values = c("springgreen4", "indianred3")) +
  labs(x = "museum open or closed", y = "count", fill = NULL,
       title = "accreditation status by museum open/closed status") +
  theme(plot.title = element_text(hjust = 0.5))
```

### script runtime

```{r}
tictoc::toc()
```













